---
title: "Week 3B Window Function"
author: "Renee Watson"
date: "2026-09-16"
output: html_document
---

# Window Function

## Approach

In this assignment, I will use an LLM-generated dataset containing daily sales for three products: Coffee, Tea, and Juice. The dataset includes daily observations from January 1 through March 31, 2026. This dataset provides a time series for multiple items, which will allow me to practice using window functions in R. My plan is to load the dataset into R and use dplyr to organize the data by item and date. I will calculate the year-to-date average, sales for each product and the six-day moving average. I will group the data by item so that the calculations are performed separately for the products.

One challenge I anticipate is making sure that the observations are arranged in the correct date order and grouped by product so that the year-to-date and six-day moving averages are calculated correctly.

Data Source: daily_product_sales_2026.csv

## Codebase

```{r}
library(dplyr)
library(slider)
```

I loaded the daily sales dataset into R and examined its structure.

```{r}
sales_data <- read.csv("daily_product_sales_2026.csv")
str(sales_data)
```

I converted the date variable to a date format and arranged the observations by product and date. This ensures that the window calculations are performed in chronological order.

```{r}
sales_data$date <- as.Date(sales_data$date)
sales_data <- sales_data %>%
  arrange(item, date)
```

## Analysis

I grouped the data by product so that the calculations would be performed separately for Coffee, Tea, and Juice. I calculated the year-to-date average which includes all sales observations for a product from the beginning of the year through the current date. I also calculated a six-day moving average, which uses the current day and the previous five days of sales.

```{r}
sales_data <- sales_data %>%
  group_by(item) %>%
  mutate(
    ytd_average = cummean(sales),
    six_day_average = slide_dbl(
      sales,
      mean,
      .before = 5,
      .complete = TRUE
    )
  ) %>%
  ungroup()

```

```{r}
sales_data %>% 
  group_by(item) %>%
  slice_head(n = 10) %>%
  ungroup()
```

## Conclusion

In this analysis, I calculated year-to-date and six-day moving averages for daily sales of Coffee, Tea, and Juice. Grouping the data by product allowed each product's averages to be calculated independently. The year-to-date average provides a cumulative measure of sales performance, while the six-day moving average focuses on more recent sales.

## AI Use

ChatGPT was used as an AI-assisted tool to generate the artificial daily sales dataset and to assist with code development, debugging, and refinement. Its suggestions and generated code were reviewed, tested, and revised as needed. I remained responsible for the submitted analysis, code, and conclusions.

Tool/model: ChatGPT/ GPT 5.6 Sol Developer: OpenAI Date accessed: September 18, 2026
